Practical Machine Learning Course Project

This document is the course project of "Practical Machine Learning" on coursera. It is built up in Rstudio, using its knitr function.

#Overview and Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). The goal of the project is to predict the manner in which they did the exercise. 

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.
#Data Processing
##Loading data and libraries 
```{r,echo=TRUE}
library(gbm)
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(gbm)
trainingurl<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingurl<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```

##Cleaning Data 
when we check the Excel file, we found a lot NA,#DIV/0! and blank values
we remove the variable that contains more than 95% of the observation to be these values
```{r,echo=TRUE}
trainingRaw<-read.csv(url(trainingurl),na.strings = c("","NA","#DIV/0!"))
testingRaw<-read.csv(url(testingurl),na.strings =c("","NA","#DIV/0!"))
columnIndex<-colSums(is.na(trainingRaw))/nrow(trainingRaw)<0.95
validtraining<-trainingRaw[,columnIndex]
validtesting<-testingRaw[,columnIndex]
dim(validtraining)
dim(validtesting)
```
Now both cleaned the training and test set has 93 variables 

Then we also need to remove the variable don't contribute too much to the accelerometer measurements
```{r,echo=TRUE}
classe<-validtraining$classe
trainremove<-grepl("^x|timestamp|window",names(validtraining))
training1<-validtraining[,!trainremove]
trainingclean<-training1[,sapply(training1,is.numeric)] ##only keep the numeric variable expect the classe
trainingclean$classe<-classe
testremove<-grepl("^x|timestamp|window",names(validtesting))
testing1<-validtesting[,!testremove]
testingclean<-testing1[,sapply(testing1,is.numeric)]
dim(trainingclean)
dim(testingclean)
```
Now, the cleaned training data set contains 19622 observations and 54 variables, while the testing data set contains 20 observations and 54 variables. 

##Silice the Data
```{r,echo=TRUE}
set.seed(1222)
inTrain<-createDataPartition(y=trainingclean$classe,p=0.6,list = FALSE)
trainingdata<-trainingclean[inTrain,]
testingdata<-trainingclean[-inTrain,]
```

##Model Selection
Decision Tree Model
```{r,echo=TRUE}
Mod1<-train(classe~.,data=trainingdata,method="rpart")
pred1<-predict(Mod1,testingdata)
confusionMatrix(pred1,as.factor(testingdata$classe))$overall[1]
```
We can see when using Decision Tree Model, the accuracy is about 66.17%

Random Forest Model
```{r,echo=TRUE}
library(parallel)
library(doParallel)
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
fitControl <- trainControl(method = "cv",number = 5,allowParallel = TRUE)
Mod2<-train(classe~.,data = trainingdata,method="rf")
pred2<-predict(Mod2,testingdata)
confusionMatrix(pred2,as.factor(testingdata$classe))$overall[1]
```

We can see when using Random Forest Model, the accuracy is about 100%

So we choose Random Forest Model

##Predictions
```{r,echo=TRUE}
pred<-predict(Mod2,testingclean)
print(pred)
```

